### 2019-05-11 code/paper.Rmd
diff in years between start of article and inscription year

---
title: "Community Data Science meets Digital Heritage: Controversies about World Heritage Sites in the Online Community of Wikipedia"
---

Load the data

```{r}
library(tidyverse)

# load the random pages
ten_k_random_wp_pages <- readRDS(here::here("data", "random_page_data_tbl.rds"))

# load the random pages revision histories
ten_k_random_wp_pages_lst <- readRDS(here::here("data", "random_page_data_lst.rds"))

# all WH sites with WP pages (n = 584)
page_data_for_all_pages <- 
  readRDS(here::here("data", 
                     "page_data_for_all_pages.rds"))

# all WH sites, with and without WP pages (n = 846)
wh_wiki_table <- 
  read_csv(here::here("data", 
                      "wh_wiki_table.csv"))

# all WH sites according to UNESCO
# from https://whc.unesco.org/en/syndication
wh_unesco <- readxl::read_excel(here::here("data",
                                           "whc-sites-2018.xls")) %>% 
  filter(category %in% c("Cultural", "Mixed"))

library("rnaturalearth")
library("rnaturalearthdata")
world <- ne_countries(scale = "medium", returnclass = "sf")

# functions used more than once:

estimate_mode <- function(x) {
  d <- density(x)
  d$x[which.max(d$y)]
}

theme_nogrid <- function (base_size = 12, base_family = "") {
  theme_bw(base_size = base_size, 
           base_family = base_family) %+replace% 
    theme(panel.grid = element_blank() )   
}
```

## Representation by country

Where are the WH sites, before looking into WP

```{r}

wh_unesco_countries <- 
wh_unesco %>% 
  separate_rows(states_name_en, convert = TRUE, sep = ",") %>% 
  mutate(country = case_when(
    states_name_en == "Bolivia (Plurinational State of)"  ~ "Bolivia",
    states_name_en == "Cabo Verde"  ~ "Republic of Cabo Verde",
    states_name_en == "Czechia"  ~ "Czech Republic",
    states_name_en == "Democratic People's Republic of Korea"  ~ "Republic of Korea",
    states_name_en == "Gambia (the)"  ~ "The Gambia",
    states_name_en == "Holy See"  ~ "Vatican",
    states_name_en == "Iran (Islamic Republic of)"  ~ "Iran",
    states_name_en == "Lao People's Democratic Republic"  ~ "Lao PDR",
    states_name_en == "Micronesia (Federated States of)"  ~ "Federated States of Micronesia",
    states_name_en == "Syrian Arab Republic"  ~ "Syria",
    states_name_en == "the former Yugoslav Republic of Macedonia"  ~ "Macedonia",
    states_name_en == "United Kingdom of Great Britain and Northern Ireland"  ~ "United Kingdom",
    states_name_en == "United Republic of Tanzania"  ~ "Tanzania",
    states_name_en == "United States of America"  ~ "United States",
    states_name_en == "Venezuela (Bolivarian Republic of)"  ~ "Venezuela",
    states_name_en == "United States of America"  ~ "United States",
    states_name_en == "Viet Nam"  ~ "Vietnam",
    states_name_en == "Republic of Moldova"  ~ "Moldova",
    TRUE ~ as.character(states_name_en)
  ))

wh_unesco_countries_count <-  
  wh_unesco_countries %>% 
  count(country)

median_number_unesco_sites_per_country <- 
  median(wh_unesco_countries_count$n)

mode_number_unesco_sites_per_country <- 
  round(estimate_mode(wh_unesco_countries_count$n), 1)

wh_unesco_country_counts_hist <- 
ggplot(wh_unesco_countries_count,
       aes(n)) +
  geom_histogram() +
  annotate("text", x = 30, y = 30, 
           label = str_glue('Median =\n{median_number_unesco_sites_per_country} WH sites/country\nMode =\n{mode_number_unesco_sites_per_country} WH sites/country'),
           size = 3) +
  theme_nogrid(8)

sf_map_data_unesco <- 
world %>% 
  left_join(wh_unesco_countries_count,
            by = c( 'name_long' = 'country')) %>% 
  select(name, n, geometry) %>% 
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  filter(name != "Antartica")

wh_country_unesco_count_map <- 
ggplot(data = sf_map_data_unesco) +
    geom_sf(aes(fill = n), lwd = 0) +
    scale_fill_viridis_c(na.value="grey90") +
  coord_sf(ylim = c(-50, 90), datum = NA) +
  theme_minimal() +
  ggtitle("Cultural Sites on the UNESCO World Heritage list",
          subtitle = "from https://whc.unesco.org/en/syndication")

wh_country_unesco_count_map +
  annotation_custom(ggplotGrob(wh_unesco_country_counts_hist), 
                    xmin = -190, 
                    xmax = -90, 
                    ymin= 10, 
                    ymax=-70)

ggsave("figures/wh_country_unesco_count_map.png")

```

What countries are most represented?

```{r}
  # get country of site from location text
  # Laos, Czech Republic, Micronesia, Zimbabwe, South Sudan, Chad, 
  # Central African Republic, Congo, Gabon, Cameroon, Nigeria, Bosnia and Herzegovina
  #  Cote d'Ivoire, Sierra Leone, Guyana, Belize, 
  country_names <-  paste(c(world$name, 
                            "United States",
                            "Czech Republic",
                             "Antigua & Barbuda",
                            "Saint Kitts and Nevis"
                            ),
                          collapse="|")

  page_data_for_all_pages_location <- 
    page_data_for_all_pages %>% 
    mutate(country = str_extract(Location, 
                                 regex(country_names, 
                                       ignore.case=TRUE))) %>% 
    mutate(country = case_when(
      country == "United States" ~ "United States of America",
      country == "Czech Republic" ~ "Czechia",
      country == "Antigua & Barbuda" ~ "Antigua and Barb.",
      country == "Saint Kitts and Nevis" ~ "St. Kitts and Nevis",
      TRUE ~ as.character(country)))

wh_wiki_table_location <- 
  wh_wiki_table %>% 
  mutate(country = str_extract(Location, 
                                 regex(country_names, 
                                       ignore.case=TRUE))) %>% 
    mutate(country = case_when(
      country == "United States" ~ "United States of America",
      country == "Czech Republic" ~ "Czechia",
      country == "Antigua & Barbuda" ~ "Antigua and Barb.",
      country == "Saint Kitts and Nevis" ~ "St. Kitts and Nevis",
      TRUE ~ as.character(country)))

```

```{r}

wh_wk_country_counts <- 
page_data_for_all_pages_location %>% 
  count(country) %>% 
  filter(!is.na(country))

theme_nogrid <- function (base_size = 12, base_family = "") {
  theme_bw(base_size = base_size, 
           base_family = base_family) %+replace% 
    theme(panel.grid = element_blank() )   
}

median_number_wp_pages_per_country <- 
  median(wh_wk_country_counts$n)

estimate_mode <- function(x) {
  d <- density(x)
  d$x[which.max(d$y)]
}

mode_number_wp_pages_per_country <- 
  round(estimate_mode(wh_wk_country_counts$n), 1)

wh_wk_country_counts_hist <- 
ggplot(wh_wk_country_counts,
       aes(n)) +
  geom_histogram() +
  annotate("text", x = 15, y = 30, 
           label = str_glue('Median = {median_number_wp_pages_per_country} WP pages/country\nMode = {mode_number_wp_pages_per_country} WP pages/country'),
           size = 2) +
  theme_nogrid(12)

wh_wk_country_counts_bar <- 
ggplot(wh_wk_country_counts,
       aes(reorder(country, n),
           n)) +
  geom_col() +
  coord_flip() +
  theme_minimal(base_size = 6) +
  xlab("")
```

Here's a map of countries showing which has the most WP pages for WH sites

```{r}
sf_map_data <- 
world %>% 
  left_join(wh_wk_country_counts,
            by = c( 'name' = 'country')) %>% 
  select(name, n, geometry) %>% 
  mutate(n = ifelse(is.na(n), 0, n)) %>% 
  filter(name != "Antartica")

wh_wk_country_count_map <- 
ggplot(data = sf_map_data) +
    geom_sf(aes(fill = n), lwd = 0) +
    scale_fill_viridis_c(na.value="grey90") +
  coord_sf(ylim = c(-50, 90), datum = NA) +
  theme_minimal() +
  ggtitle("Cultural Sites on the UNESCO World Heritage list with Wikipedia pages",
          subtitle = str_glue("The top five are {wh_wk_country_counts %>% 
                                                  arrange(-n) %>% 
                                                  slice(1:5) %>% 
                                                  pull(country) %>%
                                                  str_flatten(collapse = ', ')}"))

wh_wk_country_count_map +
  annotation_custom(ggplotGrob(wh_wk_country_counts_hist), 
                    xmin = -190, 
                    xmax = -90, 
                    ymin= 10, 
                    ymax=-70)

ggsave("figures/wh_country_wikpedia_pages_count_map.png")
```

Map of the ratio of WH sites with WP pages, per country

```{r}
wh_wiki_table_location_prop <- 
wh_wiki_table_location %>% 
  mutate(has_wp_page = ifelse(!is.na(site_page_name), 1, 0)) %>% 
  group_by(country) %>% 
  summarise(prop_sites_with_pages = sum(has_wp_page) / n()) %>% 
  filter(!is.na(country))

wh_wk_country_prop_hist <- 
ggplot(wh_wiki_table_location_prop,
       aes(prop_sites_with_pages)) +
  geom_histogram() +
  theme_nogrid(10)

wh_wk_country_all_counts <- 
wh_wiki_table_location %>% 
  count(country) %>% 
  filter(!is.na(country)) %>% 
  left_join(wh_wiki_table_location_prop)

wh_wk_country_prop_scatter <- 
ggplot(wh_wk_country_all_counts,
       aes(prop_sites_with_pages, n)) +
  geom_point(alpha = 0.5) +
  coord_fixed(0.02) +
  theme_nogrid(6)

wh_wk_country_prop_scatter_marginal <- 
ggExtra::ggMarginal(wh_wk_country_prop_scatter,
                    type = "histogram")

sf_map_data_prop <- 
world %>% 
  filter(name != "Antartica") %>% 
  left_join(wh_wiki_table_location_prop,
            by = c( 'name' = 'country')) %>% 
  select(name, 
         prop_sites_with_pages, 
         geometry)  %>% 
  mutate(prop_sites_with_pages = ifelse(is.na(prop_sites_with_pages), 0, prop_sites_with_pages))

wh_wk_country_prop_map <- 
ggplot(data = sf_map_data_prop) +
    geom_sf(aes(fill = prop_sites_with_pages), lwd = 0) +
    scale_fill_viridis_c(na.value="grey90") +
  coord_sf(ylim = c(-50, 90), datum = NA) +
  theme_minimal() +
  ggtitle("Ratio of number of Cultural Sites on the UNESCO World Heritage list\nto the number of those sites with Wikipedia pages")

wh_wk_country_prop_map +
  annotation_custom((wh_wk_country_prop_scatter_marginal), 
                    xmin = -190, 
                    xmax = -90, 
                    ymin= 10, 
                    ymax=-70)

ggsave("figures/wh_country_wikpedia_pages_ratio_map.png")
```

## Basic qualities of the content of WP articles about WH sites

```{r}
ten_k_random_wp_pages_basic_qualities <- 
ten_k_random_wp_pages %>% 
  select(page_wordcount,
         page_wikilinks_out, 
         page_cited_items_on) %>% 
  mutate(rnd_page_wikilinks_out_per_1k_words = page_wikilinks_out / page_wordcount * 1000,
         rnd_page_cited_items_on_per_1k_words = page_cited_items_on / page_wordcount * 1000,
         rnd_page_wordcount = page_wordcount) %>% 
  select(-page_wikilinks_out,
         -page_cited_items_on,
         -page_wordcount)

ten_k_random_wp_pages_basic_qualities_long <- 
  ten_k_random_wp_pages_basic_qualities %>% 
  gather(variable, value) 
```

Length: correlate with ??
number of links out/word
number of references/word: scholarly or not

```{r}
page_data_for_all_pages_basic_qualities <- 
page_data_for_all_pages %>% 
  select(page_wordcount,
         page_wikilinks_out, 
         page_cited_items_on) %>% 
  mutate(page_wikilinks_out_per_1k_words = page_wikilinks_out / page_wordcount * 1000,
         page_cited_items_on_per_1k_words = page_cited_items_on / page_wordcount * 1000) %>% 
  select(-page_wikilinks_out,
         -page_cited_items_on)

page_data_for_all_pages_basic_qualities_long <- 
page_data_for_all_pages_basic_qualities %>% 
  gather(variable, value) %>% 
  bind_rows(ten_k_random_wp_pages_basic_qualities_long) %>% 
  mutate(source = ifelse(str_detect(variable, "rnd" ), "10k Random pages", "World Heritage pages")) %>% 
  mutate(variable = str_remove(variable, "rnd_"))

label_names <- list(
  "page_wordcount"="Page word count",
  "page_wikilinks_out_per_1k_words"="Number of wikilinks\nper 1k words",
  "page_cited_items_on_per_1k_words"="Number of sources cited\nper 1k words"
)

labeller <- function(variable,value){
  return(label_names[value])
}

# density plots
  ggplot(page_data_for_all_pages_basic_qualities_long,
         aes(value)) +
  geom_density(aes(fill = source),
               alpha = 0.3) +
  scale_x_log10() +
  scale_fill_viridis_d() +
  facet_wrap( ~ variable, 
              scales = "free",
              labeller = labeller) +
  theme_minimal() +
    ggtitle("Basic qualities about Wikipedia article content")

  ggsave("figures/wh_wikpedia_pages_basic_content.png")
```

```{r}
# scatterplots
  ggplot() +
      geom_point(data = ten_k_random_wp_pages_basic_qualities,
             aes(rnd_page_wikilinks_out_per_1k_words,
                 rnd_page_cited_items_on_per_1k_words,
                 size = rnd_page_wordcount),
             colour = "grey80",
             alpha = 0.4) +
  geom_point(data = page_data_for_all_pages_basic_qualities,
             aes(page_wikilinks_out_per_1k_words,
                 page_cited_items_on_per_1k_words,
                 size = page_wordcount),
             colour = "red",
             alpha = 0.8) +
  scale_x_log10() +
  scale_y_log10() +
  scale_fill_viridis_d() +
  coord_fixed() +
  theme_minimal()

```

## Basic qualities of the consumption of WP articles about WH sites

number of links in
page_views_last_n_days_total

```{r}
library(ggrepel)
ten_k_random_wp_pages_attent <- 
ten_k_random_wp_pages %>% 
  select(page_wikilinks_in, 
         page_views_last_n_days_total)

page_data_for_all_pages_attent <- 
page_data_for_all_pages %>% 
  select(page_wikilinks_in, 
         page_views_last_n_days_total)

page_data_for_all_pages_attent_long <- 
page_data_for_all_pages_attent %>% 
  gather(variable, value) %>% 
  bind_rows(ten_k_random_wp_pages_attent %>% 
              gather(variable, value) %>% 
              mutate(variable =  str_glue('rnd_{variable}'))) %>% 
  mutate(source = ifelse(str_detect(variable, "rnd" ), "10k Random pages", 
                         "World Heritage pages")) %>% 
  mutate(variable = str_remove(variable, "rnd_"))

# density plots

label_names <- list(
  "page_wikilinks_in"="Number of wikilinks to the page",
  "page_views_last_n_days_total"="Total page views\n(last 100 days)"
)

labeller <- function(variable,value){
  return(label_names[value])
}

  ggplot(page_data_for_all_pages_attent_long,
         aes(value)) +
  geom_density(aes(fill = source),
               alpha = 0.3) +
  scale_x_log10() +
  scale_fill_viridis_d() +
  facet_wrap( ~ variable, 
              scales = "free",
              labeller = labeller) +
  theme_minimal() +
    ggtitle("Basic qualities about consumption of Wikipedia articles")

ggsave("figures/wh_wikpedia_pages_consumption.png")
```

Scatterplot showing the WP sites that get the most attention

```{r}
library(ggrepel)
page_data_for_all_pages %>% 
  select( Site,
          page_wordcount,
         page_wikilinks_in, 
         page_views_last_n_days_total)  %>% 
  ggplot(aes(page_wikilinks_in,
             page_views_last_n_days_total,
             label = str_wrap(Site, 30))) +
  geom_point(alpha = 0.2,
             aes(size = page_wordcount,
                 colour = page_wordcount)) +
    geom_text_repel(
    data          = page_data_for_all_pages %>%  
                      filter(page_views_last_n_days_total > 190000),
    segment.size  = 0.2,
    force = 50,
    segment.color = "grey50",
    direction     = "both"
  ) +
  scale_y_log10(expand = c(1, 0.05)) +
  scale_x_log10(limits = c(1, 10000)) +
  coord_fixed(0.5) +
  scale_color_viridis_c(name = "Wordcount", 
                        labels = c(20000, 15000, 10000, 5000),
                        breaks = c(20000, 15000, 10000, 5000)) +
  scale_size_continuous(name = "Wordcount", 
                        labels = c(20000, 15000, 10000, 5000),
                        breaks = c(20000, 15000, 10000, 5000)) +
   guides(color= guide_legend(), size=guide_legend()) +
  theme_minimal(base_size = 14) +
  labs(x = "Number of wikilinks to the page",
       y = "Total page views (last 100 days)") +
    ggtitle("Basic qualities about the consumption of Wikipedia articles")

ggsave("figures/wh_wikpedia_pages_consumption_scatterplot.png")
```

## Basic qualities of the editing process of WP articles about WH sites

number of edits, total, edit density (prop/words), size of edits
how many bot edits, how many reverts per page
diversity of editors

```{r}
# wh sites on wp
revision_history_page_details <- 
  tibble(revision_history_page_details = map(page_data_for_all_pages$page_info_t, 
      ~.x$revision_history_page_details)) %>% 
  mutate(Site =         page_data_for_all_pages$Site,
         Country =      page_data_for_all_pages$country,
         rh_n_editors = map_int(revision_history_page_details, ~n_distinct(.x$rh_user)),
         rh_n_edits =   map_int(revision_history_page_details, ~nrow(.x)),
         rh_user_simpson_idx = page_data_for_all_pages$rh_user_simpson_idx,
         rh_user_bot_prop = page_data_for_all_pages$rh_user_bot_prop,
         rh_revert_prop = page_data_for_all_pages$rh_revert_prop)

# random sites

ten_k_random_wp_pages_revision_history_page_details <- 
  map_df(ten_k_random_wp_pages_lst, 
         ~.x$revision_history_page_details, 
         .id = 'Site')

# very slow!
ten_k_random_wp_pages_revision_history_page_details_summary_tbl <- 
  ten_k_random_wp_pages_revision_history_page_details %>% 
  mutate(Site = as.numeric(Site)) %>% 
  group_by(Site) %>% 
  summarise(
         rh_n_editors = n_distinct(rh_user),
         rh_n_edits =   n(),
         rh_n_bot_edits = sum(bot),
         rh_n_reverts = sum(revert)) %>% 
  mutate(rh_revert_prop = rh_n_reverts / rh_n_edits,
         rh_bot_edits_prop = rh_n_bot_edits / rh_n_edits)

# plot both samples, number of edits per page
bind_rows(revision_history_page_details %>% 
            select(Site, rh_n_edits) %>% 
            mutate(Site = "wh"),
          ten_k_random_wp_pages_revision_history_page_details_summary_tbl %>% 
            select(Site, rh_n_edits) %>% 
            mutate(Site = "rand")) %>% 
  ggplot(aes(rh_n_edits,
             fill = Site)) +
  geom_density(alpha = 0.3) +
  scale_x_log10() +
  scale_fill_viridis_d(labels = c("Random", 
                               "World Heritage")) +
  labs(x = "Number of edits per page",
       fill = "Wikipedia pages") +
  theme_minimal()

ggsave("figures/wh_wikpedia_edits_per_page.png")
```

Edits per 1k words, as a measure of edit density

```{r}
revision_history_page_details_wordcount <- 
  revision_history_page_details %>% 
  bind_cols(page_data_for_all_pages %>% select(page_wordcount)) %>% 
  mutate(edits_per_1k_words = rh_n_edits / (page_wordcount / 1000))

ten_k_random_wp_pages_revision_history_page_details_summary_tbl_all_dets <- 
bind_cols(ten_k_random_wp_pages, 
          ten_k_random_wp_pages_revision_history_page_details_summary_tbl) %>% 
  mutate(edits_per_1k_words = rh_n_edits / (page_wordcount / 1000))

bind_rows(revision_history_page_details_wordcount %>% 
            select(Site, edits_per_1k_words) %>% 
            mutate(Site = "wh"),
          ten_k_random_wp_pages_revision_history_page_details_summary_tbl_all_dets %>% 
            select(Site, edits_per_1k_words) %>% 
            mutate(Site = "rand")) %>% 
  ggplot(aes(edits_per_1k_words,
             fill = Site)) +
  geom_density(alpha = 0.3) +
  #xlim(0, 100) +
  scale_x_log10() +
  scale_fill_viridis_d(labels = c("Random", 
                               "World Heritage")) +
  labs(x = "Number of edits per 1k words per page",
       fill = "Wikipedia pages") +
  theme_minimal()

ggsave("figures/wh_wikpedia_edits_density_page.png")
```

Size of edits, absolute and raw

```{r}
page_data_for_all_pages_rh_tbl <- 
map_df(page_data_for_all_pages$page_info_t, 
      ~.x$revision_history_page_details, .id = "Site")

label_names <- list(
  "mean_abs_diff"="Absolute edit sizes",
  "mean_diff"="Raw edit sizes"
)

labeller <- function(variable,value){
  return(label_names[value])
}

bind_rows(page_data_for_all_pages_rh_tbl %>% 
            select(Site, rh_diff_size) %>% 
            group_by(Site) %>% 
            summarise(mean_diff = mean((rh_diff_size)),
                      mean_abs_diff = mean(abs(rh_diff_size))) %>% 
            mutate(Site = "wh"),
          ten_k_random_wp_pages_revision_history_page_details %>% 
            select(Site, rh_diff_size) %>% 
              group_by(Site) %>% 
            summarise(mean_diff = mean((rh_diff_size)),
                      mean_abs_diff = mean(abs(rh_diff_size))) %>% 
            mutate(Site = "rand")) %>% 
  gather(variable, value, -Site) %>% 
  ggplot(aes(value,
             fill = Site)) +
  geom_density(alpha = 0.3) +
  #xlim(0, 100) +
  scale_x_log10() +
  scale_fill_viridis_d(labels = c("Random", 
                               "World Heritage")) +
  labs(x = "Average size of edits per article (bytes)",
       fill = "Wikipedia pages") +
  facet_wrap( ~ variable, labeller = labeller) +
  theme_minimal()

ggsave("figures/wh_wikpedia_edit_sizes.png")
```

How many edits made by bots? Proportion of edits per page made by bots

```{r}
bind_rows(

page_data_for_all_pages_rh_tbl %>% 
  group_by(Site) %>% 
  summarise( n_bot_edits = sum(bot),
             n_edits = n()) %>% 
  mutate(prop_bot_edits = n_bot_edits / n_edits,
           Site = "wh"),

ten_k_random_wp_pages_revision_history_page_details %>%  
  group_by(Site) %>% 
summarise( n_bot_edits = sum(bot),
           n_edits = n()) %>% 
  mutate(prop_bot_edits = n_bot_edits / n_edits,
         Site = "rand")
) %>%

  ggplot(aes(prop_bot_edits,
             fill = Site)) +
  geom_density(alpha = 0.3) +
  #xlim(0, 100) +
  scale_x_log10() +
  scale_fill_viridis_d(labels = c("Random", 
                               "World Heritage")) +
  labs(x = "Proportion of edits made by bots per article",
       fill = "Wikipedia pages") +
  theme_minimal()

ggsave("figures/wh_wikpedia_edit_sizes.png")
```

Which WH pages are most edited by bots?

```{r}
library(ggrepel)

wh_bots <- 
 tibble(revision_history_page_details = map(page_data_for_all_pages$page_info_t, 
      ~.x$revision_history_page_details))  %>% 
  mutate(Site = page_data_for_all_pages$Site) %>% 
  mutate(n_bot_edits = map_dbl(revision_history_page_details, ~sum(.x$bot)),
         n_edits = map_int(revision_history_page_details, ~nrow(.))) %>% 
  mutate(prop_bot_edits = n_bot_edits / n_edits) %>% 
  arrange(desc(prop_bot_edits))

  ggplot(wh_bots, 
         aes(n_bot_edits,
             n_edits,
             label = str_wrap(Site, 30))) +
  geom_point(alpha = 0.2,
             aes(size =   prop_bot_edits,
                 colour = prop_bot_edits)) +
    geom_text_repel(
    data            = wh_bots %>%  
                      filter(prop_bot_edits > 0.36),
    segment.size  = 0.2,
    force = 50,
    segment.color = "grey50",
    direction     = "both"
  ) +
  scale_y_log10() +
  scale_x_log10() +
  coord_fixed(0.5) +
  scale_color_viridis_c(name = "Proportion of\nedits by bots", 
                        labels = c(0, 0.1, 0.2, 0.3, 0.4),
                        breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  scale_size_continuous(name = "Proportion of\nedits by bots", 
                        labels = c(0, 0.1, 0.2, 0.3, 0.4),
                        breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
   guides(color= guide_legend(), size=guide_legend()) +
  theme_minimal(base_size = 14) +
  labs(x = "Number of edits by bots",
       y = "Total number of edits") +
    ggtitle("Bot activity on Wikipedia articles")

ggsave("figures/wh_wikpedia_pages_bot_edits_scatterplot.png")
```

which bots are most active on WH pages vs random pages?
Cydebot: category edits
Smackbot: dates in templates
ClueBot NG: vandalism

```{r}
bind_rows(
page_data_for_all_pages_rh_tbl %>% 
  filter(bot == 1) %>% 
  group_by(rh_user) %>% 
  tally(sort = TRUE) %>% 
  slice(1:10) %>% 
  mutate( Site = "World Heritage Sites"),

ten_k_random_wp_pages_revision_history_page_details %>% 
  filter(bot == 1) %>% 
  group_by(rh_user) %>% 
  tally(sort = TRUE) %>% 
    slice(1:10) %>% 
   mutate( Site = "Random pages")) %>%

  ggplot(aes(reorder(rh_user, n),
             n)) +
  geom_col() +
  theme_minimal() +
  coord_flip() +
  facet_wrap( ~ Site, scales = "free") +
  xlab("")

ggsave("figures/wh_wikpedia_what_bots.png")
```

So about vandalism? do WH sites get more than random ones?

```{r}
bind_rows(
page_data_for_all_pages_rh_tbl %>% 
  mutate(vandalism = ifelse(str_detect(rh_comment, "vandal"), 1, 0)) %>%
  group_by(Site) %>% 
  summarise( n_vandal_edits = sum(vandalism),
             n_edits = n()) %>% 
  mutate(prop_vandal_edits = n_vandal_edits / n_edits,
           Site = "World Heritage Sites"),

ten_k_random_wp_pages_revision_history_page_details %>% 
  mutate(vandalism = ifelse(str_detect(rh_comment, "vandal"), 1, 0)) %>%
  group_by(Site) %>% 
  summarise( n_vandal_edits = sum(vandalism),
             n_edits = n()) %>% 
  mutate(prop_vandal_edits = n_vandal_edits / n_edits, 
         Site = "Random pages")
) %>%

   ggplot(aes(prop_vandal_edits,
             fill = Site)) +
  geom_density(alpha = 0.3) +
  scale_x_log10() +
  scale_fill_viridis_d(labels = c("Random", 
                               "World Heritage")) +
  labs(x = "Proportion of edits relating to vandalism per article",
       fill = "Wikipedia pages") +
  theme_minimal()

ggsave("figures/wh_wikpedia_vandalism_edits.png")

```

Which WH pages have the most vandalism?

```{r}

vandal_edits <- 
 tibble(revision_history_page_details = map(page_data_for_all_pages$page_info_t, 
      ~.x$revision_history_page_details))  %>% 
  mutate(Site = page_data_for_all_pages$Site) %>% 
  mutate(n_vandal_edits = map_dbl(revision_history_page_details, 
                                  ~sum(ifelse(str_detect(.x$rh_comment, "vandalism"), 1, 0))),
         n_edits = map_int(revision_history_page_details, ~nrow(.))) %>% 
  mutate(prop_vandal_edits = n_vandal_edits / n_edits) %>% 
  arrange(desc(prop_vandal_edits))

ggplot(vandal_edits, 
         aes(n_vandal_edits,
             n_edits,
             label = str_wrap(Site, 30))) +
  geom_point(alpha = 0.2,
             aes(size =   prop_vandal_edits,
                 colour = prop_vandal_edits)) +
    geom_text_repel(
    data            = vandal_edits %>%  
                      filter(prop_vandal_edits > 0.055),
    segment.size  = 0.2,
    force = 50,
    segment.color = "grey50",
    direction     = "both"
  ) +
  scale_y_log10() +
  scale_x_log10() +
  coord_fixed(0.5) +
  scale_color_viridis_c(name = "Proportion of\nedits about vandalism", 
                        labels = c(0, 0.025, 0.05, 0.075),
                        breaks = c(0, 0.025, 0.05, 0.075)) +
  scale_size_continuous(name = "Proportion of\nedits about vandalism", 
                        labels = c(0, 0.025, 0.05, 0.075),
                        breaks = c(0, 0.025, 0.05, 0.075)) +
   guides(color= guide_legend(), size=guide_legend()) +
  theme_minimal(base_size = 14) +
  labs(x = "Number of edits about vandalism",
       y = "Total number of edits") +
    ggtitle("Editing about vandalism on Wikipedia articles")

ggsave("figures/wh_wikpedia_pages_vandal_edits_scatterplot.png")

```

Reverts

```{r}

bind_rows(
page_data_for_all_pages_rh_tbl %>% 
  mutate(revert = ifelse(revert, 1, 0)) %>%
  group_by(Site) %>% 
  summarise( n_revert_edits = sum(revert),
             n_edits = n()) %>% 
  mutate(prop_revert_edits = n_revert_edits / n_edits,
           Site = "World Heritage Sites"),

ten_k_random_wp_pages_revision_history_page_details %>% 
  mutate(revert = ifelse(revert, 1, 0)) %>%
  group_by(Site) %>% 
  summarise( n_revert_edits = sum(revert),
             n_edits = n()) %>% 
  mutate(prop_revert_edits = n_revert_edits / n_edits, 
         Site = "Random pages")
) %>%

   ggplot(aes(prop_revert_edits,
             fill = Site)) +
  geom_density(alpha = 0.3) +
  scale_x_log10() +
  scale_fill_viridis_d(labels = c("Random", 
                               "World Heritage")) +
  labs(x = "Proportion of reverting edits per article",
       fill = "Wikipedia pages") +
  theme_minimal()

```

What's the typical distribution of talk page length?

```{r}
bind_rows(

ten_k_random_wp_pages %>% 
  select(talk_page_wordcount) %>% 
  mutate(Site = "Random pages"),

page_data_for_all_pages %>% 
  select(talk_page_wordcount) %>% 
  mutate(Site = "World Heritage pages")
) %>%

  ggplot(aes(talk_page_wordcount, 
             fill = Site)) +
  geom_density(alpha = 0.3) +
   scale_fill_viridis_d(labels = c("Random", 
                               "World Heritage")) +
  labs(x = "Talk page word count",
       fill = "Wikipedia pages") +
  theme_minimal() +
  scale_x_log10(labels = scales::comma_format(accuracy = 1))

ggsave("figures/wh_wikpedia_pages_talk_page_worcount.png")
```

which WH pages have the longest talk pages?

```{r}

wh_pages_talk_page_length <- 
left_join(page_data_for_all_pages,
          revision_history_page_details_wordcount)

ggplot(wh_pages_talk_page_length,
         aes(rh_n_edits,
             rh_revert_prop,
             label = str_wrap(Site, 30))) +
  geom_point(alpha = 0.5,
             aes(size =   (talk_page_wordcount),
                 colour = (talk_page_wordcount))) +
    geom_text_repel(
    data            = wh_pages_talk_page_length %>%  
                      filter(talk_page_wordcount > 8000),
    segment.size  = 0.2,
    colour = "black",
    force = 50,
    segment.color = "grey50",
    direction     = "both"
  ) +
  scale_y_log10() +
  scale_x_log10(labels = scales::comma_format(accuracy = 1)) +
  coord_fixed(1) +
  scale_color_viridis_c(name = "Talk page word count", 
                        labels = c(5000, 10000, 15000, 20000),
                        breaks = c(5000, 10000, 15000, 20000)) +
  scale_size_continuous(name = "Talk page word count", 
                        labels = c(5000, 10000, 15000, 20000),
                        breaks = c(5000, 10000, 15000, 20000)) +
   guides(color= guide_legend(), size=guide_legend()) +
  theme_minimal(base_size = 14) +
  labs(x = "Number of edits",
       y = "Proportion of edits that are reverts") +
    ggtitle("")

ggsave("figures/wh_wikpedia_pages_talk_page_wordcount_reverts_scatterplot.png")

```

Talk page topic modelling

```{r}

```

## Locality

IP addresses

```{r}

ip_regex <- "(?(?=.*?(\\d+\\.\\d+\\.\\d+\\.\\d+).*?)(\\1|))"

get_ip_addresses <- function(x){
  re <- regexpr(
  ip_regex, 
  x, perl = TRUE)
regmatches(x, re)
}

# get locations from IP addresses
library(rgeolocate)
ip_db <- system.file("extdata","GeoLite2-Country.mmdb", package = "rgeolocate")

```

Do WH sites have more geographically diverse contributions?

Get locations for random pages.

How many IP edits on each page?

```{r}
randoms_revision_history_page_details <- 
  map_df(ten_k_random_wp_pages_lst, 
         ~.x$revision_history_page_details, 
         .id = "page")

randoms_revision_history_page_details_ip <- 
randoms_revision_history_page_details %>% 
    mutate(is_ip_address = ifelse(get_ip_addresses(rh_user) != "", TRUE, FALSE),
         page = as.numeric(page))

randoms_revision_history_page_details_ip_prop <- 
  randoms_revision_history_page_details_ip %>% 
  group_by(page) %>% 
  summarise(prop_ip_addresses = sum(is_ip_address) / n() )

ggplot(randoms_revision_history_page_details_ip_prop,
       aes(prop_ip_addresses)) +
  geom_histogram() +
  theme_minimal()
```

Overall location of IP edits for all random pages

```{r}
randoms_revision_history_page_details_ip_location <- 
randoms_revision_history_page_details_ip %>% 
  filter(is_ip_address) %>% 
  mutate(ip_location = map(rh_user, ~maxmind(.x, ip_db)))  %>% 
  unnest(ip_location)

randoms_revision_history_page_details_ip_location_tally <- 
randoms_revision_history_page_details_ip_location %>% 
  filter(!is.na(country_name)) %>% 
  mutate(country_name = fct_rev(fct_infreq(country_name))) %>% 
  group_by(country_name) %>% 
  tally(sort = TRUE) %>% 
  filter(n > 100)

ggplot(randoms_revision_history_page_details_ip_location_tally,
       aes(country_name, n)) +
  geom_col() +
  coord_flip()   +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma)
```

What's the geographic diversity of edits per random page?

```{r}
randoms_revision_history_page_details_ip_location_props_per_page <- 
randoms_revision_history_page_details_ip_location %>% 
  filter(!is.na(country_name)) %>% 
  mutate(country_name = fct_rev(fct_infreq(country_name))) %>% 
  group_by(page, country_name) %>% 
  tally() %>% 
  mutate(prop = n / sum(n)) %>% 
  filter(country_name %in% randoms_revision_history_page_details_ip_location_tally$country_name)

ggplot(randoms_revision_history_page_details_ip_location_props_per_page,
       aes(country_name, 
           prop)) +
  geom_boxplot() +
  coord_flip()

randoms_revision_history_page_details_ip_location_props_per_page_div <- 
  randoms_revision_history_page_details_ip_location_props_per_page %>% 
  group_by(page) %>% 
  summarise(shannon = vegan::diversity(n))

ggplot(randoms_revision_history_page_details_ip_location_props_per_page_div,
       aes( shannon)) +
  geom_histogram() 
```

```{r}
# analyse the countries of IP address vs country of WH site
revision_history_page_details_ip_addresses <- 
revision_history_page_details %>% 
  mutate(edits_from_ip_addresses = map(revision_history_page_details,
                          ~.x %>% 
                            filter(get_ip_addresses(rh_user) != "") %>% 
                            mutate(ip_location = map(rh_user, 
                                                     ~maxmind(.x, ip_db)))  %>% 
                            unnest(ip_location)))

revision_history_page_details_ip_addresses_df <- 
revision_history_page_details_ip_addresses %>% 
  unnest(edits_from_ip_addresses)

revision_history_page_details_ip_addresses_countries <- 
revision_history_page_details_ip_addresses_df %>% 
  select(Site, Country, country_name, rh_n_edits ) %>% 
  filter_all(all_vars(!is.na(.))) %>% 
  group_by(Site, Country, rh_n_edits) %>% 
  count(country_name) %>% 
  mutate(prop_anon_of_all_edits = n / rh_n_edits) %>% 
  mutate(internal_edits = ifelse(Country ==  country_name, 
                                 TRUE, FALSE),
         n_anon_edits_total = n) %>% 
  left_join(revision_history_page_details_ip_addresses_df) %>% 
  distinct(Site,  
           Country, 
           prop_anon_of_all_edits, 
           internal_edits,
           .keep_all = TRUE) %>% 
  group_by(Site,  Country) %>% 
  summarise(sum_prop_internal = sum(n_anon_edits_total[internal_edits]) / sum(n_anon_edits_total),
            sum_prop_external = 1 - sum_prop_internal,
            rh_n_edits = unique(rh_n_edits),
            rh_n_edits_anon = sum(n_anon_edits_total),
            prop_anon_of_all_edits = sum(n) / rh_n_edits)
```

```{r}
# Histogram of proportion of anonymous edits 
# TODO and random sites
ggplot(revision_history_page_details_ip_addresses_countries,
       aes(prop_anon_of_all_edits)) +
  geom_histogram() +
  theme_minimal()
```

```{r}
# Histogram of proportion of all anonymous edits that come from same country as WH site
ggplot(revision_history_page_details_ip_addresses_countries,
       aes(sum_prop_internal)) +
  geom_histogram() +
  theme_minimal() 
```

```{r}
ggplot(revision_history_page_details_ip_addresses_countries,
       aes(sum_prop_internal,
           prop_anon_of_all_edits)) +
  geom_point() +
  theme_minimal() +
  coord_equal()

```

```{r}
adjacency_list_3_cols <- 
revision_history_page_details_ip_addresses_df %>% 
  filter(!is.na(Country),
         !is.na(country_name)) %>% 
  group_by(Country, country_name) %>% 
  count() %>% 
  rename(to = Country,
         from = country_name,
         value = n) %>% 
  ungroup() %>% 
  mutate(from = ifelse(from == "Hashemite Kingdom of Jordan", "Jordan", from)) %>% 
  select(from, to , value) %>% 
  filter(value > 100) %>% 
  left_join( revision_history_page_details_ip_addresses_df %>% 
               select(country_name, continent_name) %>% 
               distinct(),
             by = c('from' = 'country_name')) %>% 
  arrange(continent_name,  from, to)

adjacency_list <- 
adjacency_list_3_cols%>% 
  select(-continent_name)

get_countries_in_continent <- function(x) {

  xx <- 
  revision_history_page_details_ip_addresses_df %>% 
    select(country_name, continent_name) %>% 
    filter(continent_name == x) %>% 
    pull(country_name) %>% 
    unique()

  xx[
 unique( c(adjacency_list$from,
           adjacency_list$to))
 ]
}

africa   <- get_countries_in_continent("Africa")        
asia    <- get_countries_in_continent("Asia")       
europe  <-get_countries_in_continent("Europe")  
north_america  <- get_countries_in_continent("North America")         
oceania <-  get_countries_in_continent("Oceania") 
south_america  <-  get_countries_in_continent("South America")

library(circlize)
circos.clear()
chordDiagram(adjacency_list, 
             directional = 1,
             direction.type = c("diffHeight", "arrows"),
             link.arr.type = "big.arrow", 
             diffHeight = -uh(2, "mm"),
             annotationTrack = "grid", 
             preAllocateTracks = list(track.height = max(0.4)))

# we go back to the first track and customize sector labels
circos.track(track.index = 1, panel.fun = function(x, y) {
    circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index, 
        facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5))
}, bg.border = NA) # here set bg.border to NA is important

highlight.sector(africa, track.index = 1, col = "#00FF0040", 
    text = "Africa", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(asia, track.index = 1, col = "#0000FF40", 
    text = "Asia", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(europe, track.index = 1, col = "blue", 
    text = "Europe", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(north_america, track.index = 1, col = "orange", 
    text = "North America", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(oceania, track.index = 1, col = "purple", 
    text = "Oceania", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(south_america, track.index = 1, col = "brown", 
    text = "South America", cex = 0.8, text.col = "white", niceFacing = TRUE)
```

```{r}

#-- 
circos.track(track.index = 2, panel.fun = function(x, y) {
    sector.index = get.cell.meta.data("sector.index")
    xlim = get.cell.meta.data("xlim")
    ylim = get.cell.meta.data("ylim")
    circos.text(mean(xlim), mean(ylim), sector.index, cex = 0.6, niceFacing = TRUE)
}, bg.border = NA)

#--
circos.clear()
circos.par(gap.after = rep(c(rep(1, 4), 8), 3))

chordDiagram(adjacency_list, annotationTrack = c("grid", "axis"),
    preAllocateTracks = list(
        track.height = uh(4, "mm"),
        track.margin = c(uh(4, "mm"), 0)
))

circos.track(track.index = 2, panel.fun = function(x, y) {
    sector.index = get.cell.meta.data("sector.index")
    xlim = get.cell.meta.data("xlim")
    ylim = get.cell.meta.data("ylim")
    circos.text(mean(xlim), mean(ylim), sector.index, cex = 0.6, niceFacing = TRUE)
}, bg.border = NA)

highlight.sector(africa, track.index = 1, col = "red", 
    text = "Africa", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(asia, track.index = 1, col = "green", 
    text = "Asia", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(europe, track.index = 1, col = "blue", 
    text = "Europe", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(north_america, track.index = 1, col = "orange", 
    text = "North America", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(oceania, track.index = 1, col = "purple", 
    text = "Oceania", cex = 0.8, text.col = "white", niceFacing = TRUE)

highlight.sector(south_america, track.index = 1, col = "brown", 
    text = "South America", cex = 0.8, text.col = "white", niceFacing = TRUE)

```

## Networks

editors

## Controversies

Look at distributions of variables, then say 2sd is extreme, and match those to the WH pages to find out what they are

```{r}
```

Can we detect -ve sentiment on the talk pages? Perhaps as distributions, over every 100 words per page

```{r}
library(tidytext)

```

What's the typical distribution of the proportion of edits that are reverts?

```{r}
ggplot(ten_k_random_wp_pages,
       aes(rh_revert_prop)) +
  geom_histogram() +
  scale_x_log10()

```

What are the keywords for controversy in edit messages? And how common are these warring messages?

```{r}
randoms_revision_history_page_details_ip_warring_words <- 
randoms_revision_history_page_details_ip %>% 
  mutate(warring_words = (str_detect(rh_comment, 'undid|vandalism|why|inaccurate|wrong|warning|blocked|warring|disputed|neutral|unacceptable' )))

# distribution of edits with warring words
randoms_revision_history_page_details_ip_warring_words_prop <- 
randoms_revision_history_page_details_ip_warring_words %>% 
  group_by(page) %>% 
  summarise(prop_warring_edits = sum(warring_words) / n())

ggplot(randoms_revision_history_page_details_ip_warring_words_prop,
         aes(prop_warring_edits)) +
  geom_histogram() +
  scale_x_log10()
```

Is there are correlation between the proportion of warring words and talk page length?

```{r}
# some interesting groupings
randoms_revision_history_page_details_ip_warring_words_prop %>% 
  left_join(ten_k_random_wp_pages %>% 
  mutate(page = 1:nrow(.))) %>% 
  filter(prop_warring_edits > 0) %>% 
  ggplot(aes(prop_warring_edits, 
             talk_page_wordcount ,
             size = page_wordcount)) +
  geom_point(alpha = 0.2)
```

Is there are correlation between the proportion of warring words and revert frequency?

```{r}
randoms_revision_history_page_details_ip_warring_words_prop_revert_prop <- 
randoms_revision_history_page_details_ip_warring_words_prop %>% 
  left_join(ten_k_random_wp_pages %>% 
  mutate(page = 1:nrow(.))) %>% 
  filter(prop_warring_edits != 0,
         rh_revert_prop != 0)

lm(rh_revert_prop ~ prop_warring_edits, 
   data = randoms_revision_history_page_details_ip_warring_words_prop_revert_prop) %>% summary

ggplot(randoms_revision_history_page_details_ip_warring_words_prop_revert_prop,
         aes(prop_warring_edits, 
             rh_revert_prop)) +
  geom_point()
```

time between edits

```{r}
library(lubridate)
randoms_revision_history_page_details_ip_time_diffs <- 
randoms_revision_history_page_details_ip %>% 
  group_nest(page) %>% 
  mutate(time_diff = map(data, ~lag(.x$rh_date) - .x$rh_date)) %>% 
  unnest(time_diff) %>% 
  mutate(time_diff_days = as.numeric(time_diff / 60 / 60 / 24)) %>% 
  mutate(time_diff_hours = as.numeric(time_diff / 60 / 60 )) %>% 
  mutate(less_than_one_month = ifelse(time_diff_hours < 744, TRUE, FALSE)) %>% 
  mutate(less_than_one_week = ifelse(time_diff_hours < 168, TRUE, FALSE)) %>% 
  bind_cols(randoms_revision_history_page_details_ip)

rle_tbl <- 
tibble(lths = rle(randoms_revision_history_page_details_ip_time_diffs$rh_user)$lengths,
vals = rle(randoms_revision_history_page_details_ip_time_diffs$rh_user)$values)

rle_tbl2 <- tibble(
  rh_user = rep(rle_tbl$vals, rle_tbl$lths),
  lths2 = rep(rle_tbl$lths, rle_tbl$lths)
)

randoms_revision_history_page_details_ip_time_diffs %>% 
  filter(!is.na(time_diff_hours)) %>% 
  mutate(quick_three_edits = time_diff_hours + lag(time_diff_hours) + lag(time_diff_hours, 2) + lag(time_diff_hours, 3))

randoms_revision_history_page_details_ip_time_diffs_quick <- 
randoms_revision_history_page_details_ip_time_diffs %>% 
 nest(-page) %>%  
 mutate(edits_3 = map(data, ~.x %>% 
                              filter(less_than_one_week) %>% 
                              filter(n() > 4 ))) %>% 
  unnest(edits_3)

randoms_revision_history_page_details_ip_time_diffs %>% 
  bind_cols(rle_tbl2) %>% 
  filter(lths2 == 1) %>% 
  filter(time_diff_hours < 24)

ggplot(randoms_revision_history_page_details_ip_time_diffs,
       aes(time_diff_days)) +
  geom_histogram()

```

## Temporal patterns

Page ages: are WH pages younger than the typical WP page? What's their relationship to their inscription year?

```{r}
# get first and last edit, and inscription year
library(lubridate)
page_data_for_all_pages_rh_tbl_edit_start_end_span <- 
page_data_for_all_pages_rh_tbl %>% 
  group_by(Site) %>% 
  summarise(first_edit_date = min(rh_date),
            last_edit_date = max(rh_date),
            edit_span = last_edit_date - first_edit_date) %>% 
  mutate(Site = as.numeric(Site)) %>% 
  arrange(Site) %>% 
  bind_cols(page_data_for_all_pages) %>% 
  mutate(inscription_year = as.Date(ISOdate(Year_num, 1, 1))) %>% 
  mutate(diff_first_edit_inscription_year = as.numeric(ymd_hms(first_edit_date) - as.POSIXlt(inscription_year)) / 365 )

ggplot(page_data_for_all_pages_rh_tbl_edit_start_end_span, 
       aes(diff_first_edit_inscription_year)) +
  geom_histogram() +
  theme_minimal(base_size = 12)  +
  geom_vline(xintercept = 0, colour = "red") +
  ylim(0, 60) +
  annotate("text", x = -10, y = 55, label = "Page appeared\nbefore inscription\n←", size = 10) +
  annotate("text", x = 10,  y = 55, label = "Page appeared\nafter inscription\n→", size = 10)

```

time series of edits: breakouts relating to events? monthly seems like a good level of aggregation. What can we do for spike detection?

```{r}
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 130) %>%   # explore this value
  ggplot() +
  geom_line(aes(the_m,
             n)) +
  facet_wrap( ~ page, scales = "free")
```

Spike detection

for randoms, what even is an anomaly?

```{r}
# devtools::install_github("hrbrmstr/AnomalyDetection")
library(AnomalyDetection)
randoms_revision_history_page_details_ip_location_monthly <- 
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 150) # explore this value

randoms_revision_history_page_details_ip_location_monthly_anom <- 
  randoms_revision_history_page_details_ip_location_monthly %>% 
  ungroup %>% 
  group_nest(page) %>% 
  mutate(anomalies = map(data, ~ad_ts(.x, 
                                      max_anoms = 0.2, 
                                      alpha = 0.005,
                                      threshold = "p95"))) %>% 
  unnest(anomalies)

ggplot() +
  geom_line(
    data=randoms_revision_history_page_details_ip_location_monthly %>% 
      filter(page %in% randoms_revision_history_page_details_ip_location_monthly_anom$page), 
    aes(the_m, n), 
    size=0.125, color="lightslategray")  +
  geom_point(
    data=randoms_revision_history_page_details_ip_location_monthly_anom, 
    aes(timestamp, anoms), 
    color="#cb181d", alpha=1/3) +
  scale_x_datetime(date_labels="%b\n%Y") +
  facet_wrap( ~ page, scales = "free_y")

```

for WH sites

```{r}

n <- 12 # months of edit data
revision_history_page_details_ip_addresses_df_monthly_edits <- 
revision_history_page_details_ip_addresses_df  %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(Site, the_m) %>% 
  tally() %>% 
  filter(n() >  n ) # n months worth of data - explore this value

# distribution of edits per month
ggplot(revision_history_page_details_ip_addresses_df_monthly_edits,
       aes(n)) +
  geom_histogram() +
  scale_x_log10()

# subset so we get only sites with > n1 edits/month, and at least 3 months worth of edits

n1 <- 7  # edits per month
n2 <- 20 # months worth of edits
revision_history_page_details_ip_addresses_df_monthly_edits_10 <- 
revision_history_page_details_ip_addresses_df_monthly_edits %>% 
  filter(max(n) >= n1) %>% 
  group_by(Site) %>% 
  filter(n() >= n2)

```

```{r}

# detect anomalies
revision_history_page_details_ip_addresses_df_monthly_edits_anoms <- 
  revision_history_page_details_ip_addresses_df_monthly_edits_10 %>% 
  ungroup %>% 
  group_nest(Site) %>% 
  mutate(anomalies = map(data, ~ad_ts(.x, 
                                      max_anoms = 0.2, 
                                      alpha = 0.005,
                                      threshold = "p95"))) %>% 
  unnest(anomalies) %>% 
  group_by(Site) %>% 
  filter(max(anoms) >= 10) # where the anomaly is n edits/month or more

ggplot() +
  geom_line(
    data=revision_history_page_details_ip_addresses_df_monthly_edits_10 %>% 
      filter(Site %in% revision_history_page_details_ip_addresses_df_monthly_edits_anoms$Site), 
    aes(the_m, n), 
    size=0.125, color="lightslategray")  +
  geom_point(
    data=revision_history_page_details_ip_addresses_df_monthly_edits_anoms, 
    aes(timestamp, anoms), 
    color="#cb181d", alpha=1/3) +
  scale_x_datetime(date_labels="%b\n%Y") +
  facet_wrap( ~ Site, scales = "free")
```

## Technicity

bots

What do bots do?

Do people undo bot actions often?

  - length of articles http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.451.4208&rep=rep1&type=pdf 
  - edit history to identify: diversity of contributions, 
  - reverts in edits for spotting contentious topics https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0038869 http://www.contropedia.net/  http://wwm.phy.bme.hu/ 
    - vandalism traces in edit history
  - talk page length
  - number of links in each page
  - page views: https://www.sciencedirect.com/science/article/abs/pii/S0006320716301240 
  - timing of edits: political events? Brexit?
    - citations in articles: https://arxiv.org/abs/0705.2106
  - are the citations scholarly? Archaeological? Meskell?
    - citations to these articles: https://www.tandfonline.com/doi/full/10.1080/0194262X.2016.1206052 
  - network analysis of editors, clusters of editors
  - sentiment analysis of talk pages
  - geolocation of anon editors by IP address
  - bots and technicity https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0171774 but see https://dl.acm.org/citation.cfm?id=3134684

  - Compare to 10000 random Wikipedia sites
  - Compare to WH sites in media & academic literature